Analysis of pleiotropy of eQTLs

### 2023-01-19 presentation_230120_internal_seminar/intern_sem_230120.tex
ENH

        \begin{itemize}
            \item Association test of effect allele on case individuals
            \item High-throughput genetics: 3000 unique loci for over 250 disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories
        \end{itemize}
%
        \vfill
%          However, how do we know if these variants are active in a tissue,
%         or whether these variants false positives due to linkage disequilibrium,
%         or how these variants act molecularly
        Questions:
%
        \begin{itemize}
            \item Correlation between variants. Which is the causal variant?
            \item Which cell types important for disease?
            \item In which tissues are these variants active?
        \end{itemize}
%
        \vfill
%
        Questions partially addressed by looking at eQTLs in different tissues

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Expression quantitative trait loci (eQTLs)}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}

            \end{column}
        \end{columns}

@@

@@

        \begin{enumerate}
            \item Large scale gene and tissue annotation of variants of complex traits using GWAS/eQTL colocalization analysis
            \item To identify pleiotropic eQTLs and eQTL-containing genomic regions
            \item To characterize the properties of pleiotropic eQTLs
            \item To share colocalizations with experimentalists
        \end{enumerate}
    \end{frame}

@@

@@

@@

@@

@@

        \begin{itemize}
            \item Example eQTLs involved in four GWAS categories in different cytobands
            \item The gene marker is the eQTL gene with the highest PubMed publication count
        \end{itemize}

% full size table is table
        \begin{table}[!tbp]
            \centering
            \footnotesize
            \csvreader[separator=tab,
            tabular=crclcp{0.4\textwidth},
            head,
            table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Known eQTL gene & \bfseries GWAS Categories\\\hline,
            ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
                {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
            \vspace{15pt}
        \end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Concluding remarks}

@@

        \begin{itemize}
            \item A web portail to share data
            \item Causality of eQTL genes on GWAS (Looking for experts ;-) )
        \end{itemize}

>>>>

        \begin{itemize}
            \item Association test of effect allele on case individuals
            \item High-throughput genetics: 3000 unique loci for over 250 disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories
        \end{itemize}
%
        \vfill

%          However, how do we know if these variants are active in a tissue,
%         or whether these variants false positives due to linkage disequilibrium,
%         or how these variants act molecularly
        Questions:
%
        \begin{itemize}
            \item Correlation between variants. Which is the causal variant?
            \item Which cell types important for disease?
            \item In which tissues are these variants active?
        \end{itemize}
%
        \vfill
%
        Questions partially addressed by looking at eQTLs in different tissues

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Expression quantitative trait loci (eQTLs)}
%eQTLs are variants that modify gene expression in a given tissue.
%        eQTLs give a potential mechanism (the gene target) and a tissue.
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}

            \end{column}
        \end{columns}
%
        \vfill
%
%        To my knowledge, it is unknown where are the most pleiotropice Eqtls in the genome.
%        This is important because this would give hints to putatitve pleiotropi genes (unknown or unknown) that are regulated by these pleiotropic eQTLs?
        Pleiotropy of eQTLs in genome is currently unknown

@@

        \begin{enumerate}
            \item To look for pleiotropic eQTLs in the genome
            \item To characterise their genomic regions, target genes and the cause of their pleiotropy
        \end{enumerate}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Strategy}

        \begin{enumerate}
            \item Annotate eQTLs with GWAS traits using colocalization analysis
            \item Annotate GWAS traits in categories
            \item Aggregate trait categories of eQTLs: Pleiotropic eQTLs
            \item Aggregate pleiotropic eQTLs in genome regions
            \item Analyse pleiotropic eQTLs and regions
        \end{enumerate}
    \end{frame}

@@

@@

%In this project, I have carried a large colocalization analysis between ~300 GWAS traits and 127 eQTL studies belonging to different tissues and mainly primary cell types.
%In addition, by colocalizing statistically eQTLs and GWAS, we can to some extent, break, linkage disequilibrium

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Web portail}

%These data can be useful to interpret disease variants
%        Therefore I have developed a small web portail

        TODO

@@

%        To evaluate the coherence of the colocalization data, I have computed distances between traits based on the Spearman correlation of the betas of the eqtl-gene-tissue

@@

%        In this heatmap, we can see that similar traits cluster together
%One difficulty for analysing pleiotropy is to distinguish similar traits for really different traits.
%In the previous heatmap, we have seen that traits cluster around coherent categories
%To look for real pleiotropic eQTLs I have manually annotated traits in different categories, such as cardiovascular disease, allergic, autoimmune, ....

@@

        \begin{itemize}
            \item Example eQTLs involved in four GWAS categories in different cytobands
            \item The representative gene is the eQTL gene with the highest PubMed publication count
        \end{itemize}

% full size table is table
        \begin{table}[!tbp]
            \centering
            \footnotesize
            \csvreader[separator=tab,
            tabular=crclcp{0.4\textwidth},
            head,
            table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Known eQTL gene & \bfseries GWAS Categories\\\hline,
            ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
                {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
            \vspace{15pt}\label{tab:pleitropic_variants}
        \end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Conclusion}

        \begin{itemize}
            \item A list of pleiotrtA large set of causal GWAS variants annotated with target genes and tissues
            \item A list of pleiotropic variants and regions
        \end{itemize}

@@

        \begin{itemize}
            \item Causality of eQTL genes on GWAS
            \item Study loci in more
        \end{itemize}
